---
title: "COVID-19 in Bulgaria"
author: '[Metodi Simeonov](https://www.linkedin.com/in/mesime/){target="_blank"}'
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    source_code: embed
runtime: shiny
---

``` {css}

a {
    color: #ffffff;
    text-decoration: none;
}

.navbar-inverse .navbar-nav>.active>a,.navbar-inverse .navbar-nav>.active>a:hover,.navbar-inverse .navbar-nav>.active>a:focus {
    color: #ffffff;
    background-color: black;
}

.navbar-inverse .navbar-nav>li>a:hover,.navbar-inverse .navbar-nav>li>a:focus {
    color: #ffffff;
    background-color: black;
}

.navbar-inverse .navbar-nav>.active>a,.navbar-inverse .navbar-nav>.active>a:hover,.navbar-inverse .navbar-nav>.active>a:focus {
    color: #ffffff;
    background-color: black;
}

body {
    font-family: "Source Sans Pro",Calibri,Candara,Arial,sans-serif;
    font-size: 15px;
    line-height: 1.42857143;
    color: #ffffff;
}

.navbar-dark .navbar-brand, .navbar.navbar-inverse .navbar-brand {
    color: black;
}

.chart-title {
  background-color: black;
  color: white;
}

.chart-stage {
  background-color: black;
}

.section.sidebar {
  background-color: black;
}

.navbar.navbar-inverse {
  background-color: #18bc9c;
  box-shadow: 3px 3px 5px black;
}

.nav-tabs .nav-link.active, .nav-tabs>li>a.active {
    background-color:#18bc9c;
}

.navbar-dark .navbar-brand, .navbar.navbar-inverse .navbar-brand {
  color:white;
}

.navbar.navbar-inverse ul.nav.navbar-nav>li.active>a {
  color: #18bc9c;
}

.dashboard-column {
  background-color: #18bc9c;
}

.section {
  background-color: #18bc9c;
}

body {
  background-color:#18bc9c;
}

```

```{r setup, include=FALSE}

# Core
library(tidyverse)
library(remotes)
# devtools::install_github("RamiKrispin/coronavirus",
#                          upgrade = c("always"),
#                          quiet   = FALSE,
#                          force   = TRUE)
library(coronavirus)

# Interactivity
library(flexdashboard)
library(shiny)

# Map & Visuals
library(leaflet)
library(leafpop)
library(plotly)
library(ggplot2)
library(grid)
library(gridExtra)
library(viridis)

`%>%` <- magrittr::`%>%`

# Set colors - https://www.w3.org/TR/css-color-3/#svg-color 
# recovered_color <- "deepskyblue"

confirmed_color <- "#012833" # rgba(1, 74, 95, 0.97)
active_color    <- "#012833"
death_color     <- "#1f0601"

df                     <- read_rds("df.rds")
df_daily               <- read_rds("df_daily.rds")
daily_confirmed_deaths <- read_rds("daily_confirmed_deaths.rds")
daily_confirmed        <- read_rds("daily_confirmed.rds")
cv_data_for_plot       <- read_rds("cv_data_for_plot.rds")
cv_data_for_plot.split <- read_rds("cv_data_for_plot.split.rds")
joined_df              <- read_rds("joined_df.rds")
mortality_rate         <- read_rds("mortality_rate.rds")

```

About
=====================================

***The COVID-19 Dashboard***

This dashboard provides an overview of the Coronavirus (COVID-19 / SARS-CoV-2) epidemic for Bulgaria and its surrounding countries. 

It is built with R using R Markdown, flexdashboard, and Shiny framework. The first version of the dashboard was adapted from [**Rami Krispin**](https://ramikrispin.github.io){target="_blank"}, courtesy of [**Antoine Soetewey**](https://github.com/AntoineSoetewey){target="_blank"}.

***How to use it***

All plots are made with Plotly and thus are interactive. Vaccination / Mortality tab contains static render graph. 

Double click to zoom into specific time windows or to select / de-select specific countries from the legend. 

Automatic updates are performed daily at 11:00 AM EET.

Application code including automation and data wrangling scripts can be found on my [***Github***](https://github.com/Met0o/COVID-19-Bulgaria-v2)

***Packages***

-   Dashboard interface - [***flexdashboard***](https://rmarkdown.rstudio.com/flexdashboard/) and [***Shiny***](https://shiny.rstudio.com/)
-   Visualization - [***plotly***](https://plot.ly/r/)
-   Data manipulation - [***dplyr***](https://dplyr.tidyverse.org/), [***tidyr***](https://tidyr.tidyverse.org/), and [***purrr***](https://purrr.tidyverse.org/)
-   Mapping - [***leaflet***](https://rstudio.github.io/leaflet/) and [***leafpop***](https://github.com/r-spatial/leafpop)

***Data***

The input data for this dashboard is the dataset available from the [***`{coronavirus}`***](https://github.com/RamiKrispin/coronavirus){target="_blank"} R package.

The raw data is pulled from Johns Hopkins University Center for Systems Science and Engineering (JHU CCSE) Coronavirus [***repository***](https://github.com/RamiKrispin/coronavirus-csv){target="_blank"}.

Summary Tab
=====================================

## Row {data-width="350"}

### Confirmed cases {.value-box}

```{r}

renderValueBox({

valueBox(
  value = paste(format(sum(df$confirmed), big.mark = ","), "", sep = " "),
  caption = "Total Confirmed Cases",
  icon = "fas fa-user-md",
  color = confirmed_color
        )
      }
    )

```

### Mortality {.value-box}

```{r}

renderValueBox({

valueBox(
  value = paste(format(sum(df$death, na.rm = TRUE), big.mark = ","), " (",
    round(100 * sum(df$death, na.rm = TRUE) / sum(df$confirmed), 1),
    "%)",
    sep = ""
  ),
  caption = "Deceased / Mortality Rate",
  icon = "fa-cross",
  color = death_color
        )
      }
    )

```

## Column {data-width="400"}

### **Daily Cumulative Cases by Type** (Bulgaria only)

```{r}

output$plotly_3 <- renderPlotly(expr = {

plot_ly(data = df_daily) %>%
  add_trace(
    x = ~date,
    y = ~confirmed_cum,
    type = "scatter",
    mode = "lines+markers",
    name = "Confirmed",
    line = list(color = active_color),
    marker = list(color = active_color)
  ) %>%
  add_trace(
    x = ~date,
    y = ~death_cum,
    type = "scatter",
    mode = "lines+markers",
    name = "Deceased",
    line = list(color = death_color),
    marker = list(color = death_color)
  ) %>%
  add_annotations(
    x = as.Date("2020-03-11"),
    y = 3,
    text = paste("First Death"),
    xref = "x",
    yref = "y",
    arrowhead = 5,
    arrowhead = 3,
    arrowsize = 1,
    showarrow = TRUE,
    ax = -90,
    ay = -90
  ) %>%
  add_annotations(
    x = as.Date("2020-03-13"),
    y = 14,
    text = paste(
      "Lockdown"
    ),
    xref = "x",
    yref = "y",
    arrowhead = 5,
    arrowhead = 3,
    arrowsize = 1,
    showarrow = TRUE,
    ax = -10,
    ay = -90
  ) %>%
  layout(
    paper_bgcolor = "#ffffff",  # change this hex to modify background color
    plot_bgcolor  = "#ffffff",  # change this hex to modify background color
    title = "",
    yaxis = list(title = ""),
    xaxis = list(title = ""),
    legend = list(x = 0.1, y = 0.9),
    hovermode = "compare"
    )
  }
) %>% 
  bindCache(input$x, input$y)

plotlyOutput(outputId = "plotly_3")

```

## Column {data-width="400"}

### **Regional Mortality**

```{r daily_summary}

output$plotly_2 <- renderPlotly(expr = {

daily_confirmed_deaths %>%
  plot_ly(colors = "BrBG") %>%
  add_trace(
    x = ~date,
    y = ~Bulgaria,
    marker = list(color = "#9119b8"),
    line = list(color = "#9119b8"),
    type = "scatter",
    mode = "lines+markers",
    name = "Bulgaria"
  ) %>%
  add_trace(
    x = ~date,
    y = ~Romania,
    marker = list(color = "#07108a"),
    line = list(color = "#07108a"),
    type = "scatter",
    mode = "lines+markers",
    name = "Romania"
  ) %>%
  add_trace(
    x = ~date,
    y = ~`North Macedonia`,
    marker = list(color = "#044a00"),
    line = list(color = "#044a00"),
    type = "scatter",
    mode = "lines+markers",
    name = "N.Macedonia"
  ) %>%
    plotly::add_trace(
    x = ~date,
    y = ~Serbia,
    marker = list(color = "#4a0000"),
    line = list(color = "#4a0000"),
    type = "scatter",
    mode = "lines+markers",
    name = "Serbia"
  ) %>%
  add_trace(
    x = ~date,
    y = ~Greece,
    marker = list(color = "#00a6a3"),
    line = list(color = "#00a6a3"),
    type = "scatter",
    mode = "lines+markers",
    name = "Greece"
  ) %>%
  layout(
    paper_bgcolor = "#ffffff",  # change this hex to modify background color
    plot_bgcolor  = "#ffffff",  # change this hex to modify background color
    title  = "",
    legend = list(x = 1, y = 0.9),
    yaxis  = list(title = ""),
    xaxis  = list(title = ""),
    hovermode = "compare",
    margin = list(
      b = 10,
      t = 10,
      pad = 2
      )
    )
  }
) %>% 
  bindCache(input$x, input$y)
  
plotlyOutput(outputId = "plotly_2")
  
```

Comparison by Country
=====================================

## Row {data-width="350"}

### **Daily New Confirmed Cases**

```{r}
 
output$plotly_1 <- renderPlotly(expr = {

daily_confirmed %>%
  plot_ly(colors = "Accent") %>%
  add_trace(
    x = ~date,
    y = ~Bulgaria,
    marker = list(color = "#9119b8"),
    line = list(color = "#9119b8"),
    type = "scatter",
    mode = "lines+markers",
    name = "Bulgaria"
  ) %>%
  add_trace(
    x = ~date,
    y = ~Romania,
    marker = list(color = "#07108a"),
    line = list(color = "#07108a"),
    type = "scatter",
    mode = "lines+markers",
    name = "Romania"
  ) %>%
  add_trace(
    x = ~date,
    y = ~`North Macedonia`,
    marker = list(color = "#044a00"),
    line = list(color = "#044a00"),
    type = "scatter",
    mode = "lines+markers",
    name = "N.Macedonia"
  ) %>%
  add_trace(
    x = ~date,
    y = ~Serbia,
    marker = list(color = "#4a0000"),
    line = list(color = "#4a0000"),
    type = "scatter",
    mode = "lines+markers",
    name = "Serbia"
  ) %>%
  add_trace(
    x = ~date,
    y = ~Greece,
    marker = list(color = "#00a6a3"),
    line = list(color = "#00a6a3"),
    type = "scatter",
    mode = "lines+markers",
    name = "Greece"
  ) %>%
  layout(
    title  = "",
    legend = list(x = 1, y = 0.9),
    yaxis  = list(title = ""),
    xaxis  = list(title = ""),
    hovermode = "compare",
    margin = list(
      b = 10,
      t = 10,
      pad = 2
      )
    )
  }
) %>% 
  bindCache(input$x, input$y)

plotlyOutput(outputId = "plotly_1")

```

## Row

### **Mortality Ratio** (Countries with More than 100,000 Cases)

```{r}

output$plotly_4 <- renderPlotly(expr = {

mortality_rate %>% 
  plot_ly(x = ~ round(100 * death_rate, 1),
                  size  = ~  log(confirmed),
                  sizes = c(5, 70),
                  #colors = "Dark2", 
                  type = 'scatter', mode = 'markers',
                  color = ~ country,
                  marker = list(sizemode = 'diameter' , opacity = 0.5),
                  hoverinfo = 'text',
                  text = ~paste("", country, 
                                " Confirmed Cases: ", confirmed,
                                " Death Rate: ",  paste(round(100 * death_rate, 1), "%", sep = ""))) %>%
   layout(title = "",
     yaxis = list(title = "", ticksuffix = "%"),
                 xaxis = list(title = "", ticksuffix = "%",
                              dtick = 1,
                              tick0 = 0),
                 hovermode = "compare")
              }
            )

plotlyOutput(outputId = "plotly_4")

```

Vaccination vs Mortality
=====================================

## Column {data-width="400"}

### **Vaccination Rate vs Mortality Rate in European Countries since March 2020**

```{r}

renderPlot({

g_mid <- ggplot(joined_df, aes(x = 1, y = country)) + 
    geom_text(aes(label = country))+
    geom_segment(aes(x = 0.94, xend = 0.96, yend  = country )) +
    geom_segment(aes(x = 1.04, xend = 1.065, yend = country)) +
    ggtitle("") +
    ylab(NULL) +
    scale_x_continuous(expand = c(0, 0), limits = c(0.94, 1.065)) +
    
    theme(axis.title.x = element_blank(), 
          axis.title.y = element_blank(), 
          axis.text.x  = element_blank(),
          axis.text.y  = element_blank(),
          axis.ticks.x = element_blank(),
          axis.ticks.y = element_blank(),
          panel.grid   = element_blank(),
          panel.background = element_blank(),
          plot.margin  = unit(c(1, -1, 1, -1), "mm"))
    
low  <- colorRampPalette(c('red', 'yellow' ))
mid  <- colorRampPalette(c("blue", "lightblue"))
high <- colorRampPalette(c('lightgreen', 'darkgreen'))

theme_update(plot.title = element_text(hjust = 0.5))


g1 <- ggplot(data = joined_df, 
             aes(x = country, y = as.numeric(fully_vaccinated_ratio), 
                 fill = fully_vaccinated_ratio, 
                 label = scales::percent(fully_vaccinated_ratio, accuracy = 0.01))) +
    xlab(NULL) +
    geom_bar(stat = "identity", alpha = 0.7, width = 0.6) + 
    
    geom_text(size = 3, position = position_stack(vjust = -0.3)) +
    
    scale_colour_viridis(direction = -1) +
    scale_fill_viridis(direction = -1) +
    
    ggtitle("Vaccination % of Adult Population") +
    theme(axis.title.x = element_blank(), 
          axis.title.y = element_blank(), 
          axis.text.x  = element_blank(),
          axis.text.y  = element_blank(),
          axis.ticks.x = element_blank(),
          axis.ticks.y = element_blank(),
          
          panel.grid.minor = element_blank(),
          panel.grid.major = element_blank(),
          
          legend.position  = "none",
          
          plot.margin  = unit(c(1, -1, 1, 0), "mm")) +
    scale_y_reverse() +
    coord_flip()


g2 <- ggplot(data = joined_df, 
             aes(x = country, y = as.numeric(death_rate), 
                 fill = death_rate, 
                 label = scales::percent(death_rate, accuracy = 0.01))) + 
    xlab(NULL) +
    geom_bar(stat = "identity", alpha = 0.7, width = 0.6) + 
    
    geom_text(size = 3, position = position_stack(vjust = 1.2)) +
    
    scale_colour_viridis_c(direction = -1) +
    scale_fill_viridis_c(direction = -1) +
    
    ggtitle("Mortality Rate") +
    theme(axis.title.x = element_blank(), 
          axis.title.y = element_blank(), 
          axis.text.x  = element_blank(),
          axis.text.y  = element_blank(),
          axis.ticks.x = element_blank(),
          axis.ticks.y = element_blank(),
          
          panel.grid.minor = element_blank(),
          panel.grid.major = element_blank(),
          
          legend.position  = "none",
          
          plot.margin  = unit(c(1, 0, 1, -1), "mm")) +
    coord_flip()


gg1    <- ggplot_gtable(ggplot_build(g1))
gg2    <- ggplot_gtable(ggplot_build(g2))
gg_mid <- ggplot_gtable(ggplot_build(g_mid))

grid.arrange(gg1, gg_mid, gg2, ncol = 3, widths = c(4/9, 2.8/9, 4/9))
  
})

```

World Map
=====================================

## Column {data-width="400"}

### **World map of cases** (*use + and - icons to zoom in/out*)

```{r}

# Map tab created by Art Steinmetz

pal <- colorFactor(c("#4682B4", "#451208"), domain = c("confirmed", "death"))

map_object <- leaflet() %>% addProviderTiles(providers$Stamen.Toner)

names(cv_data_for_plot.split) %>%
  walk(function(df) {
    map_object <<- map_object %>%
      addCircleMarkers(
        data = cv_data_for_plot.split[[df]],
        lng = ~long, lat = ~lat,
        color = ~ pal(type),
        stroke = FALSE,
        fillOpacity = 0.6,
        radius = ~log_cases,
        popup = popupTable(cv_data_for_plot.split[[df]],
          feature.id = FALSE,
          row.numbers = FALSE,
          zcol = c("type", "cases", "country", "province")
        ),
        group = df,
        labelOptions = labelOptions(
          noHide = FALSE,
          direction = "auto"
      )
    )
  }
)

renderLeaflet({

map_object %>%
  addLayersControl(
    overlayGroups = names(cv_data_for_plot.split),
    options = layersControlOptions(collapsed = FALSE)
    )
  }
)

```
